---
title: "covid 19"
author: "Diego Martinez"
date: "6/27/22"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was encoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
State of a patient must be dead or alive - we adjust that as well.   
  
# Question 1 - Is dying independent of sex

$H_0$:There is no correlation between the two. The sex of a person does not determine whether they have a higher/lower likelihood of dying or not. You have the same likelihood of death based off of gender. Dying is independent. 

$H_A$: There is a correlation between the two. The sex of a person does determine whether they ave a higher/lower likelihood of dying or not. You don't have the same likelihood of death based off of gender. Dying is dependent. 


Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

# Methods 


Since both our variables are categorical. We will use a bar plot with a ch-squared test and a fisher exact test. We can't use a T.test or an AOV because it's Cat vs 
Cat. Which is why we're using a different method. 


### Graphical Descriptive Results

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

  Based off of the first bar chart, men seem to be more susceptible to Coronacirus than women. Men have a higher death rate compared to women. Women have a higher alive rate compared to men.  Approximately, the frequency for men living is 750 lower than that of women. Based off of the second bar chart, women have a 55% survival rate while men have a 40% survival rate. Women have a higher survival rate by 15%. The death percentage for men is higher, as expected. Men death rate is at 60% while women death rate is at 45%. Men death rate is higher by 15%. 


  
## Numerical results


## Inferential results 




## Conclusions

