---
title: "covid 19"
author: "Diego Martinez"
date: "6/27/22"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was encoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
State of a patient must be dead or alive - we adjust that as well.   
  
# Question 1 - Is dying independent of sex

$H_0$:There is no correlation between the two. The sex of a person does not determine whether they have a higher/lower likelihood of dying or not. You have the same likelihood of death based off of gender. Dying is independent. 

$H_A$: There is a correlation between the two. The sex of a person does determine whether they ave a higher/lower likelihood of dying or not. You don't have the same likelihood of death based off of gender. Dying is dependent. 


Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

# Methods 


Since both our variables are categorical. We will use a bar plot with a ch-squared test and a fisher exact test. We can't use a T.test or an AOV because it's Cat vs 
Cat. Which is why we're using a different method. 


### Graphical Descriptive Results

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

  Based off of the first bar chart, men seem to be more susceptible to Coronacirus than women. Men have a higher death rate compared to women. Women have a higher alive rate compared to men.  Approximately, the frequency for men living is 750 lower than that of women. Based off of the second bar chart, women have a 55% survival rate while men have a 40% survival rate. Women have a higher survival rate by 15%. The death percentage for men is higher, as expected. Men death rate is at 60% while women death rate is at 45%. Men death rate is higher by 15%. 


  
## Numerical  Descriptive results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
The data shows that 36 women  out of 100 people died out in the dead group. While 64 men out of 100 people died out in the dead group. More men are dying than women. Off of the percentages, women have a 98.7% survival rate while men ave a 97% survival rate. Women have a higher survival rate by 1.7%. Men have a higher death rate by 1.6% because 3 - 1.4 = 1.6. 


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The null hypothesis has to be rejected because the p-value is .001. The p-value is not greater than .05 so the likelihood of getting these exact results is way too low. The fisher exact table is also significant in showing that we reject the null hypothesis because the p-value is .0009. Both of our p-values for the chi-square test and the fisher exact test are lower than .05, so we reject the null hypotheses.  

## Conclusions

